---
title: "Exploratory data analysis I"
subtitle: "Lecture 5"
date: "2025-09-09"
format: 
  revealjs: 
    output-file: 05-exploratory-data-analysis-I-slides.html
    footer: "[🔗 sta199-f25.github.io](https://sta199-f25.github.io/)"
    theme: slides.scss
    transition: fade
    slide-number: true
    logo: images/logo.png
    toc: false
  html: 
    code-link: true
---

```{r}
#| label: setup
#| include: false
ggplot2::theme_set(ggplot2::theme_gray(base_size = 24))
todays_ae <- ""
```

# Warm-up

## While you wait: Participate 📱💻 {.smaller}

::: {.columns}

::: {.column width="80%"}

::: wooclap
Suppose you have a dataset `df` with 100 rows and 5 columns: `x1`, `x2`, `x3`, `x4`, and `x5`. `x1` is a categorical variable with levels `a` and `b`. You run the following code:

```{r}
#| eval: false
x |>
  filter(x1 == "a") |>
  select(x1, x2, x5)
```

The resulting data frame will have:

::: wooclap-choices
- 3 columns, 50 rows 
- 3 columns, 100 rows
- 3 columns, can't tell how many rows
- 5 columns, 100 rows
- 5 columns, can't tell how many rows
:::

:::

:::

::: {.column width="20%"}

{{< include _wooclap-column.qmd >}}

:::

::: 

## Announcements {.smaller}

**Labs:**

- Submit PDF on Gradescope by the end of lab session 

- Make regular commits and push `.qmd` and PDF to GitHub

- Graded _primarily_ for attendance, participation, collaboration, and effort primarily

- Feedback provided for correctness

## Announcements {.xsmall}

**Homework:** HW 1 due Sunday 11:59pm

- Part 1: Feedback from AI
  - No Gradescope submission necessary
  - Make regular commits and push `.qmd` and PDF to GitHub
  - Immediate feedback from AI, no grading
  - **Important:** Don't forget to set homework and question number in the app when requesting feedback

- Part 2: Feedback from humans
  - Submit PDF on Gradescope by the deadline
  - Make regular commits and push `.qmd` and PDF to GitHub
  - Graded for correctness, feedback provided within ~week
  - **Important:** Don't forget to select pages corresponding to each question on Gradescope

# Exploratory data analysis

## Packages

-   For the data: [**usdata**](https://openintrostat.github.io/usdata/)

```{r}
library(usdata)
```

-   For the analysis: [**tidyverse**](https://www.tidyverse.org/packages/), [**scales**](https://scales.r-lib.org), and [**ggthemes**](https://jrnold.github.io/ggthemes/)

```{r}
#| message: false
library(tidyverse)
library(scales)
library(ggthemes)
```

## Data: `gerrymander` {.smaller}

```{r}
gerrymander
```

## What is gerrymandering?

<https://www.washingtonpost.com/business/wonkblog/gerrymandering-explained/2016/04/21/e447f5c2-07fe-11e6-bfed-ef65dff5970d_video.html>

## Participate 📱💻 

::: {.columns}

::: {.column width="75%"}

::: wooclap

You are given a new dataset to analyze. What are some of the first things you would do to get to know the data?

:::

:::

::: {.column width="25%"}

{{< include _wooclap-column.qmd >}}

:::

::: 

## Data: `gerrymander` {.smaller}

```{r}
glimpse(gerrymander)
```

## Data: `gerrymander` {.smaller}

-   Rows: Congressional districts

-   Columns:

    -   Congressional district and state

    -   2016 election: winning party, % for Clinton, % for Trump, whether a Democrat won the House election, name of election winner

    -   2018 election: winning party, whether a Democrat won the 2018 House election

    -   Whether a Democrat flipped the seat in the 2018 election

    -   Prevalence of gerrymandering: low, mid, and high

## Variable types: `district`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | [categorical, ID]{.fragment}       |
| `last_name`  |                                    |
| `first_name` |                                    |
| `party16`    |                                    |
| `clinton16`  |                                    |
| `trump16`    |                                    |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(district)
```

:::
:::

## Variable types: `last_name`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | [categorical, ID]{.fragment}       |
| `first_name` |                                    |
| `party16`    |                                    |
| `clinton16`  |                                    |
| `trump16`    |                                    |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(last_name)
```

:::
:::

## Variable types: `first_name`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | [categorical, ID]{.fragment}       |
| `party16`    |                                    |
| `clinton16`  |                                    |
| `trump16`    |                                    |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(first_name)
```

:::
:::

## Variable types: `party16`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | [categorical]{.fragment}           |
| `clinton16`  |                                    |
| `trump16`    |                                    |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(party16)
```

:::
:::

## Variable types: `clinton16`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | [numerical, continuous]{.fragment} |
| `trump16`    |                                    |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(clinton16)
```

:::
:::

## Variable types: `trump16`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | [numerical, continuous]{.fragment} |
| `dem16`      |                                    |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(trump16)
```

:::
:::

## Variable types: `dem16`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | [categorical]{.fragment}           |
| `state`      |                                    |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(dem16)
```

:::
:::

## Variable types: `state`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | categorical                        |
| `state`      | [categorical]{.fragment}           |
| `party18`    |                                    |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(state)
```

:::
:::

## Variable types: `party18`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | categorical                        |
| `state`      | categorical                        |
| `party18`    | [categorical]{.fragment}           |
| `dem18`      |                                    |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(party18)
```

:::
:::

## Variable types: `dem18`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | categorical                        |
| `state`      | categorical                        |
| `party18`    | categorical                        |
| `dem18`      | [categorical]{.fragment}           |
| `flip18`     |                                    |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(dem18)
```

:::
:::

## Variable types: `flip18`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | categorical                        |
| `state`      | categorical                        |
| `party18`    | categorical                        |
| `dem18`      | categorical                        |
| `flip18`     | [categorical]{.fragment}           |
| `gerry`      |                                    |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(flip18)
```

:::
:::

## Variable types: `gerry`

::: columns

::: {.column .xsmall}

| Variable     | Type                               |
|--------------|------------------------------------|
| `district`   | categorical, ID                    |
| `last_name`  | categorical, ID                    |
| `first_name` | categorical, ID                    |
| `party16`    | categorical                        |
| `clinton16`  | numerical, continuous              |
| `trump16`    | numerical, continuous              |
| `dem16`      | categorical                        |
| `state`      | categorical                        |
| `party18`    | categorical                        |
| `dem18`      | categorical                        |
| `flip18`     | categorical                        |
| `gerry`      | [categorical, ordinal]{.fragment}  |

:::

::: {.column .small}

```{r}
gerrymander |>
  select(gerry)
```

:::
:::


# Univariate analysis

## Univariate analysis

Analyzing a single variable:

::: incremental
-   Numerical: histogram, box plot, density plot, etc.

-   Categorical: bar plot, pie chart, etc.
:::

## Histogram - Step 1

```{r}
ggplot(gerrymander)
```

## Histogram - Step 2

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Histogram - Step 3

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram()
```

## Participate 📱💻 {.smaller}

::: {.columns}

::: {.column width="80%"}

::: wooclap
Which of the following histograms has the most appropriate binwidth for visualizing the distribution of `trump16`?

```{r}
#| layout-ncol: 4
#| fig-asp: 1
#| echo: false
#| fig-width: 5
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 1) +
  labs(title = "O Binwidth = 1")
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 5) +
  labs(title = "O Binwidth = 5")
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 10) +
  labs(title = "O Binwidth = 10")
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 25) +
  labs(title = "O Binwidth = 25")
```

:::

:::

::: {.column width="20%"}

{{< include _wooclap-column.qmd >}}

:::

::: 

## Histogram - Step 4

```{r}
#| code-line-numbers: "3"
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 5) +
  scale_x_continuous(labels = label_percent(scale = 1))
```

## Histogram - Step 5 {.smaller}

```{r}
#| code-line-numbers: "4-9"
ggplot(gerrymander, aes(x = trump16)) +
  geom_histogram(binwidth = 5) +
  scale_x_continuous(labels = label_percent(scale = 1)) +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = "Count"
  )
```

## Box plot - Step 1

```{r}
ggplot(gerrymander)
```

## Box plot - Step 2

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Box plot - Step 3

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_boxplot()
```

## Box plot - Alternative Step 2 + 3

```{r}
ggplot(gerrymander, aes(y = trump16)) +
  geom_boxplot()
```

## Box plot - Step 4

```{r}
#| code-line-numbers: "3-8"
ggplot(gerrymander, aes(x = trump16)) +
  geom_boxplot() +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = NULL
  )
```

## Density plot - Step 1

```{r}
ggplot(gerrymander)
```

## Density plot - Step 2

```{r}
ggplot(gerrymander, aes(x = trump16))
```

## Density plot - Step 3

```{r}
#| code-line-numbers: "2"
ggplot(gerrymander, aes(x = trump16)) +
  geom_density()
```

## Density plot - Step 4

```{r}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick")
```

## Density plot - Step 5

```{r}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1")
```

## Density plot - Step 6

```{r}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = 0.5)
```

## Density plot - Step 7

```{r}
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = 0.5, linewidth = 1)
```

## Density plot - Step 8 {.smaller}

```{r}
#| code-line-numbers: "3-8"
ggplot(gerrymander, aes(x = trump16)) +
  geom_density(color = "firebrick", fill = "firebrick1", alpha = 0.5, linewidth = 2) +
  labs(
    title = "Percent of vote received by Trump in 2016 Presidential Election",
    subtitle = "From each Congressional District",
    x = "Percent of vote",
    y = "Density"
  )
```

## Summary statistics

```{r}
gerrymander |>
  summarize(
    mean = mean(trump16),
    median = median(trump16)
  )
```

## Distribution of votes for Trump in the 2016 election {.smaller}

::: question
Describe the distribution of percent of vote received by Trump in 2016 Presidential Election from Congressional Districts.
:::

-   Shape: [The distribution of votes for Trump in the 2016 election from Congressional Districts is **unimodal and left-skewed**.]{.fragment}

-   Center: [The percent of vote received by Trump in the 2016 Presidential Election from a **typical** Congressional Districts is 48.7%.]{.fragment}

-   Spread: [In the **middle 50%** of Congressional Districts, 34.8% to 58.1% of voters voted for Trump in the 2016 Presidential Election.]{.fragment}

-   Unusual observations: [-]{.fragment}

# Bivariate analysis

## Bivariate analysis

Analyzing the relationship between two variables:

::: incremental
-   Numerical + numerical: scatterplot

-   Numerical + categorical: side-by-side box plots, violin plots, etc.

-   Categorical + categorical: stacked bar plots

-   Using an aesthetic (e.g., fill, color, shape, etc.) or facets to represent the second variable in any plot
:::

## Side-by-side box plots

```{r}
#| code-line-numbers: "|2|3-6|8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    y = gerry
    )
  ) +
  geom_boxplot()
```

## Summary statistics

```{r}
gerrymander |>
  summarize(
    min = min(trump16),
    q25 = quantile(trump16, 0.25),
    median = median(trump16),
    q75 = quantile(trump16, 0.75),
    max = max(trump16),
  )
```

## Participate 📱💻 {.smaller}

::: {.columns}

::: {.column width="80%"}

::: wooclap
What goes in the `[blank]` in the code below to do the following step for each level of `gerry`?

::: {.columns}
::: {.column width="60%"}

```{r}
#| eval: false
gerrymander |>
  # [blank]
  summarize(
    min = min(trump16),
    q25 = quantile(trump16, 0.25),
    median = median(trump16),
    q75 = quantile(trump16, 0.75),
    max = max(trump16),
  )
```

:::
::: {.column width="40%"}

::: wooclap-choices
- `filter(gerry)`
- `group_by(gerry)`
- `mutate(gerry)`
- `select(gerry)`
:::

:::
::: 

:::

:::

::: {.column width="20%"}

{{< include _wooclap-column.qmd >}}

:::

::: 

## Grouped summary statistics

```{r}
gerrymander |>
  group_by(gerry) |>
  summarize(
    min = min(trump16),
    q25 = quantile(trump16, 0.25),
    median = median(trump16),
    q75 = quantile(trump16, 0.75),
    max = max(trump16),
  )
```

## Density plots

```{r}
#| code-line-numbers: "|4|5|8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    color = gerry
    )
  ) +
  geom_density()
```

## Filled density plots

```{r}
#| code-line-numbers: "6"
#| output-location: column
ggplot(
  gerrymander, 
  aes(
    x = trump16, 
    color = gerry,
    fill = gerry
    )
  ) +
  geom_density()
```

## Better filled density plots

```{r}
#| code-line-numbers: "5"
ggplot(
  gerrymander, 
  aes(x = trump16, color = gerry, fill = gerry)
  ) +
  geom_density(alpha = 0.5)
```

## Better colors

```{r}
#| code-line-numbers: "7-8"
#| output-location: column
ggplot(
  gerrymander, 
  aes(x = trump16, color = gerry, fill = gerry)
  ) +
  geom_density(alpha = 0.5) +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Violin plots {.scrollable}

```{r}
#| code-line-numbers: "5"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Multiple geoms {.scrollable}

```{r}
#| code-line-numbers: "6"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_point() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Multiple geoms {.scrollable}

```{r}
#| code-line-numbers: "6"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_jitter() +
  scale_color_colorblind() +
  scale_fill_colorblind()
```

## Remove legend {.scrollable}

```{r}
#| code-line-numbers: "9"
ggplot(
  gerrymander, 
  aes(x = trump16, y = gerry, color = gerry)
  ) +
  geom_violin() +
  geom_jitter() +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  theme(legend.position = "none")
```

# Multivariate analysis

## Multivariate analysis {.smaller}

Analyzing the relationship between multiple variables:

::: incremental
-   In general, one variable is identified as the **outcome** of interest

-   The remaining variables are **predictors** or **explanatory variables**

-   Plots for exploring multivariate relationships are the same as those for bivariate relationships, but **conditional** on one or more variables

    -   Conditioning can be done via faceting or aesthetic mappings (e.g., scatterplot of `y` vs. `x1`, colored by `x2`, faceted by `x3`)

-   Summary statistics for exploring multivariate relationships are the same as those for bivariate relationships, but **conditional** on one or more variables

    -   Conditioning can be done via grouping (e.g., correlation between `y` and `x1`, grouped by levels of `x2` and `x3`)
:::

# Application exercise

## `{r} todays_ae` {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits by the end of class.
:::
