---
title: "Exploratory data analysis II"
subtitle: "Lecture 6"
date: "2025-09-11"
format: 
  revealjs: 
    output-file: 06-exploratory-data-analysis-II-slides.html
    footer: "[🔗 sta199-f25.github.io](https://sta199-f25.github.io/)"
    theme: slides.scss
    transition: fade
    slide-number: true
    logo: images/logo.png
    toc: false
  html: 
    code-link: true
---

```{r}
#| label: setup
#| include: false
ggplot2::theme_set(ggplot2::theme_gray(base_size = 24))
todays_ae <- "ae-04-gerrymander-explore-II"
```

# Warm-up

## While you wait: Participate 📱💻 {.xsmall}

::: {.columns}

::: {.column width="85%"}

::: wooclap
Which of the following is **false** about the following plot and the code that produced it?

```{r}
#| echo: false
#| warning: false
#| out-width: 100%
#| fig-width: 7
#| fig-asp: 0.25
library(tidyverse)
library(palmerpenguins)
library(ggthemes)
penguins |>
  ggplot(aes(x = species, y = flipper_length_mm, color = island, fill = island)) +
  geom_boxplot(alpha = 0.5) +
  scale_color_colorblind() +
  scale_fill_colorblind() +
  scale_y_continuous(breaks = seq(170, 230, by = 20)) +
  labs(
    title = NULL,
    x = "Species",
    y = "Flipper length (mm)",
    color = "Island", fill = "Island"
  ) +
  theme_minimal() +
  theme(
    legend.position = "top",
    plot.margin = margin(t = 0, r = 0, b = 5, l = 5, unit = "pt")
  )
```

:::

:::

::: {.column width="15%"}

{{< include _wooclap-column.qmd >}}

:::

::: 

::: wooclap
::: wooclap-choices
- There are no Chinstrap or Gentoo penguins on Torgersen Island.
- `legend.position = "bottom"` is set in the `theme()` layer.
- The same variable is mapped to both `color` and `fill`.
- `group_by(species)` is used to create the boxplots.
- A Biscoe island penguin with a flipper length of 190 mm must be an Adélie.
:::
:::

## Reminder: Code style and readability {.smaller}

::: {.incremental .xsmall}
-   Plots should include an informative title, axes and legends should have human-readable labels, and careful consideration should be given to aesthetic choices.

-   Code should follow the [tidyverse style](https://style.tidyverse.org/) ([style.tidyverse.org](https://style.tidyverse.org/)) Particularly,

    -   space before and line breaks after each `+` when building a `ggplot`
    -   space before and line breaks after each `|>` in a data transformation pipeline
    -   code should be properly indented
    -   spaces around `=` signs and spaces after commas

-   All code should be visible in the PDF output, i.e., should not run off the page on the PDF.
    Long lines that run off the page should be split across multiple lines with line breaks.
    **Tip:** Haikus not novellas when writing code!

-   Whydowecareaboutthestyleandreadabilityofyourcode? [$\rightarrow$ Why do we care about the style and readability of your code?]{.fragment}

-   Je voudrais un cafe [$\rightarrow$ Je voudrais un café]{.fragment}
:::

# `gerrymander` 

## Packages {.smaller}

-   For the data: [**usdata**](https://openintrostat.github.io/usdata/)

```{r}
library(usdata)
```

-   For the analysis: [**tidyverse**](https://www.tidyverse.org/packages/) and [**ggthemes**](https://jrnold.github.io/ggthemes/)

```{r}
library(tidyverse)
library(ggthemes)
```

## From the AE {.smaller}

::: question
Is a Congressional District more likely to have high prevalence of gerrymandering if a Democrat flipped the seat in the 2018 election? (`flip18` = 1: Democrat flipped the seat, 0: No flip, -1: Republican flipped the seat.)
:::

::: {.columns}
::: {.column}

```{r}
ggplot(
  gerrymander, 
  aes(x = flip18, fill = gerry)
  ) +
  geom_bar(position = "fill") +
  scale_fill_colorblind()
```

:::
::: {.column}

```{r}
gerrymander |>
  count(flip18, gerry) |>
  group_by(flip18) |>
  mutate(prop = n / sum(n))
```

:::
:::

# `group_by()`, `summarize()`, `count()`

## Spot the difference {.smaller}

::: question
What does `group_by()` do?
:::

::: columns
::: column

```{r}
gerrymander |>
  count(flip18, gerry)
```

:::
::: column

```{r}
gerrymander |>
  count(flip18, gerry) |>
  group_by(flip18)
```

:::
:::

## Let's simplify! {.smaller}

::: question
What does `group_by()` do in the following pipeline?
:::

::: columns
::: column

```{r}
gerrymander |>
  summarize(mean_trump16 = mean(trump16))
```

:::
::: column

```{r}
gerrymander |>
  group_by(state) |>
  summarize(mean_trump16 = mean(trump16))
```

:::
:::

## `group_by()` {.smaller}

-   Group by converts a data frame to a grouped data frame, where subsequent operations are performed once per group

-   `ungroup()` removes grouping

::: columns
::: column

```{r}
gerrymander |>
  select(state, district, party16, party18) |>
  group_by(state)
```

:::
::: {.column .fragment}

```{r}
#| code-line-numbers: "4"
gerrymander |>
  select(state, district, party16, party18) |>
  group_by(state) |>
  ungroup()
```

:::
:::

## `group_by() |> summarize()` {.smaller}

A common pipeline is `group_by()` and then `summarize()` to **calculate** summary statistics for each group:

```{r}
gerrymander |>
  group_by(state) |>
  summarize(
    mean_trump16 = mean(trump16),
    median_trump16 = median(trump16)
  )
```

## `group_by() |> summarize()` {.smaller}

This pipeline can also be used to **count** number of observations for each group:

```{r}
gerrymander |>
  group_by(state) |>
  summarize(n = n())
```

## `summarize()` {.smaller}

```{r}
#| eval: false
... |>
  summarize(
    name_of_summary_statistic = summary_function(variable)
  )
```

. . .

-   `name_of_summary_statistic`: Anything you want to call it!
    -   Recommendation: Keep it short and evocative
-   `summary_function()`:
    -   `n()`: number of observations
    -   `mean()`: mean
    -   `median()`: median
    -   `sd()`: standard deviation
    -   `min()`: minimum
    -   `max()`: maximum
    -   `IQR()`: inter-quartile range
    -   `quantile()`: quantile (e.g., 0.25, 0.75, etc.)

## Spot the difference {.smaller}

::: question
What's the difference between the following two pipelines?
:::

::: columns
::: column
```{r}
gerrymander |>
  group_by(state) |>
  summarize(n = n())
```
:::

::: column
```{r}
gerrymander |>
  count(state)
```
:::
:::

## `count()`

::: columns
::: {.column width="40%"}
``` r
... |>
  count(variable)
```
:::

::: {.column width="60%"}

```{r}
#| eval: false
... |>
  count(variable1, variable2)
```

:::
:::

-   Count the number of observations in each level of variable(s)

-   Place the counts in a variable called `n`

## Participate 📱💻 {.xsmall}

::: {.columns}

::: {.column width="80%"}

::: wooclap
How would you write the following pipeline with `count()` instead?

```{r}
#| output-location: column
gerrymander |>
  group_by(state) |>
  summarize(n = n()) |>
  arrange(desc(n))
```

::: wooclap-choices
- `gerrymander |> arrange(state) |> count()`
- `gerrymander |> count(state) |> arrange(desc(n))`
- `gerrymander |> count(state) |> sort(n)`
- `gerrymander |> count(state, sort = TRUE)`
:::

:::

:::

::: {.column width="20%"}

{{< include _wooclap-column.qmd >}}

:::

::: 


# `mutate()`

## Flip the question

::: question
Is a Congressional District more likely to have high prevalence of gerrymandering if a Democrat flipped the seat in the 2018 election?
:::

vs.

::: question
Is a Congressional District more likely to be flipped to a Democratic seat if it has high or low prevalence of gerrymandering?
:::

## What's going on? {.smaller}

::: {.question .xsmall}
The following code should produce a visualization that answers the question "Is a Congressional District more likely to be flipped to a Democratic seat if it has high or low prevalence of gerrymandering?" However, it produces a `warning` and an unexpected plot.
What's going on?
:::

```{r}
#| label: fill-warning
#| warning: true
ggplot(gerrymander, aes(x = gerry, fill = flip18)) +
  geom_bar(position = "fill")
```

## Another glimpse at `gerrymander` {.smaller}

```{r}
glimpse(gerrymander)
```

## `mutate()`

-   We want to use `flip18` as a categorical variable

-   But it's stored as a numeric

-   So we need to change its type first, before we can use it as a categorical variable

-   The `mutate()` function transforms (*mutates*) a data frame by creating a new column or updating an existing one

## `mutate()` in action {.smaller}

You can create a new variable with `mutate()`:

```{r}
#| code-line-numbers: "|2|3"
gerrymander |>
  mutate(flip18_cat = as.factor(flip18)) |>
  relocate(district, flip18, flip18_cat) # relocate to the beginning for easier viewing
```

## Revisit the plot {.smaller}

::: question
Is a Congressional District more likely to be flipped to a Democratic seat if it has high or low prevalence of gerrymandering?
:::

```{r}
#| output-location: column
#| code-line-numbers: "|2|3"
gerrymander |>
  mutate(flip18_cat = as.factor(flip18)) |>
  ggplot(aes(x = gerry, fill = flip18_cat)) +
  geom_bar(position = "fill") +
  scale_fill_colorblind()
```

## `mutate()` and overwrite

You can overwrite an existing variable with `mutate()`:

```{r}
gerrymander |>
  mutate(flip18 = as.factor(flip18)) |>
  select(district, flip18)
```

## `mutate()` and `if_else()` {.smaller}

Use `mutate()` with `if_else()` to recode with an either/or logic: 

> If `party16` is "D", recode it as "Democrat", otherwise recode it as "Republican".

```{r}
#| code-line-numbers: "|2"
gerrymander |>
  mutate(party16_expanded = if_else(party16 == "D", "Democrat", "Republican")) |>
  select(district, party16, party16_expanded)
```

## `mutate()` and `case_when()` {.smaller}

Use `mutate()` with `case_when()` to recode with a more complex logic:

> If `flip18` is 1, recode it as "Democrat flipped", if `flip18` is 0, recode it as "No flip", and if `flip18` is -1, recode it as "Republican flipped".

```{r}
gerrymander |>
  mutate(
    flip18_expanded = case_when(
      flip18 == 1 ~ "Democrat flipped",
      flip18 == 0 ~ "No flip",
      flip18 == -1 ~ "Republican flipped"
    )
  ) |>
  select(district, flip18, flip18_expanded) |>
  group_by(flip18) |> # group by flip type
  slice_head(n = 1) # show top row per group
```

## `mutate()` and store {.smaller}

If you want to keep your changes, you need to store the data frame after `mutate()`:

::: {.columns}
::: {.column}

```{r}
#| error: true
gerrymander |>
  mutate(
    p16 = if_else(party16 == "D", "Dem", "Rep")
  ) |>
  select(district, p16) |>
  slice_head(n = 3) # show top 3 rows

gerrymander |>
  select(district, p16)
```

:::
::: {.column .fragment}

```{r}
#| error: true
gerrymander <- gerrymander |>
  mutate(
    p16 = if_else(party16 == "D", "Dem", "Rep")
  ) |>
  select(district, p16)

gerrymander |>
  select(district, p16)  |>
  slice_head(n = 3) # show top 3 rows
```

:::
:::

# Application exercise

## `{r} todays_ae` {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits by the end of class.
:::

## Recap: aesthetic mappings

-   **Local** aesthetic mappings for a given `geom`

-   **Global** aesthetic mappings for all `geom`s
