---
title: "Linear models with a single predictor"
subtitle: "Lecture 16"
date: "2025-10-23"
format: 
  revealjs: 
    output-file: 16-linear-model-single-predictor-slides.html
    footer: "[🔗 sta199-f25.github.io](https://sta199-f25.github.io/)"
    theme: slides.scss
    transition: fade
    slide-number: true
    logo: images/logo.png
    pdf-separate-fragments: true
    toc: false
  html: 
    code-link: true
filters: 
  - ../remove-fmt-skip.lua
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false
ggplot2::theme_set(ggplot2::theme_gray(base_size = 24))
todays_ae <- "ae-13-modeling-penguins"
```

# Warm-up

## While you wait: Participate 📱💻 {.xsmall}

:::::: columns
:::: {.column width="70%"}
::: wooclap
Play the game a few times and report your score: smallest absolute difference between yuor guess and the actual correlation.

<https://www.rossmanchance.com/applets/2021/guesscorrelation/GuessCorrelation.html>

:::
::::

::: {.column width="30%"}
{{< include _wooclap-column.qmd >}}
:::
::::::

## Announcements {.smaller}

-   Peer evaluation 2 due on tonight at 11:59 pm

-   ...

# Linear regression with a single predictor

## Packages

```{r}
library(tidyverse)
library(tidymodels)
library(fivethirtyeight)
```

## Data prep

-   Rename Rotten Tomatoes columns as `critics` and `audience`
-   Rename the dataset as `movie_scores`

```{r}
#| label: data-prep
movie_scores <- fandango |>
  rename(
    critics = rottentomatoes,
    audience = rottentomatoes_user
  )
```

## Data overview

```{r}
#| label: data-overview
movie_scores |>
  select(critics, audience)
```

## Data visualization

```{r}
#| label: scatterplot
ggplot(movie_scores, aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) +
  labs(
    x = "Critics Score",
    y = "Audience Score"
  )
```

## Regression model {#regression-model-1}

A **regression model** is a function that describes the relationship between the outcome, $Y$, and the predictor, $X$.

$$
\begin{aligned} Y &= \color{black}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{black}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{black}{\boldsymbol{\mu_{Y|X}}} + \epsilon \end{aligned}
$$

## Regression model

::::: columns
::: {.column width="30%"}

$$
\begin{aligned} Y &= \color{#325b74}{\textbf{Model}} + \text{Error} \\[8pt]
&= \color{#325b74}{\mathbf{f(X)}} + \epsilon \\[8pt]
&= \color{#325b74}{\boldsymbol{\mu_{Y|X}}} + \epsilon 
\end{aligned}
$$

:::

::: {.column width="70%"}

```{r}
#| echo: false
#| message: false
m <- lm(audience ~ critics, data = movie_scores)
ggplot(data = movie_scores, mapping = aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "#325b74", se = FALSE, linewidth = 1.5) +
  labs(x = "X", y = "Y") +
  theme_minimal() +
  theme(
    axis.text = element_blank(),
    axis.ticks.x = element_blank(),
    axis.ticks.y = element_blank()
  )
```

:::
:::::

## Simple linear regression {.smaller}

Use **simple linear regression** to model the relationship between a quantitative outcome ($Y$) and a single quantitative predictor ($X$): $$\Large{Y = \beta_0 + \beta_1 X + \epsilon}$$

::: incremental
-   $\beta_1$: True slope of the relationship between $X$ and $Y$
-   $\beta_0$: True intercept of the relationship between $X$ and $Y$
-   $\epsilon$: Error (residual)
:::

## Simple linear regression

$$ \Large{\hat{Y} = b_0 + b_1 X} $$

-   $b_1$: Estimated slope of the relationship between $X$ and $Y$
-   $b_0$: Estimated intercept of the relationship between $X$ and $Y$
-   No error term!

## Choosing values for $b_1$ and $b_0$

```{r}
#| echo: false
#| message: false
ggplot(movie_scores, aes(x = critics, y = audience)) +
  geom_point(alpha = 0.4) +
  geom_abline(
    intercept = 32.3155,
    slope = 0.5187,
    color = "#325b74",
    linewidth = 1.5
  ) +
  geom_abline(intercept = 25, slope = 0.7, color = "gray") +
  geom_abline(intercept = 21, slope = 0.9, color = "gray") +
  geom_abline(intercept = 35, slope = 0.3, color = "gray") +
  labs(x = "Critics Score", y = "Audience Score")
```

## Residuals

```{r}
#| message: false
#| echo: false
#| fig-align: center
ggplot(movie_scores, aes(x = critics, y = audience)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "#325b74", se = FALSE, linewidth = 1.5) +
  geom_segment(
    aes(x = critics, xend = critics, y = audience, yend = predict(m)),
    color = "steel blue"
  ) +
  labs(x = "Critics Score", y = "Audience Score") +
  theme(legend.position = "none")
```

$$ \text{residual} = \text{observed} - \text{predicted} = y - \hat{y} $$

## Least squares line {.smaller}

-   The residual for the $i^{th}$ observation is

$$ e_i = \text{observed} - \text{predicted} = y_i - \hat{y}_i $$

-   The **sum of squared** residuals is

$$ e^2_1 + e^2_2 + \dots + e^2_n $$

-   The **least squares line** is the one that **minimizes the sum of squared residuals**

## Least squares line

```{r}
movies_fit <- linear_reg() |>
  fit(audience ~ critics, data = movie_scores)

tidy(movies_fit)
```

# Slope and intercept

## Properties of least squares regression

::: incremental
-   The regression line goes through the center of mass point (the coordinates corresponding to average $X$ and average $Y$): $b_0 = \bar{Y} - b_1~\bar{X}$

-   Slope has the same sign as the correlation coefficient: $b_1 = r \frac{s_Y}{s_X}$

-   Sum of the residuals is zero: $\sum_{i = 1}^n \epsilon_i = 0$

-   Residuals and $X$ values are uncorrelated
:::

## Participate 📱💻 {.xsmall}

:::::: columns
:::: {.column width="70%"}

::: wooclap
The slope of the model for predicting audience score from critics score is 0.519.
Which of the following is the best interpretation of this value?

$$\widehat{\text{audience}} = 32.3 + 0.519 \times \text{critics}$$

::: wooclap-choices
- For every one point increase in the critics score, the audience score goes up by 0.519 points, on average.
- For every one point increase in the critics score, we expect the audience score to be higher by 0.519 points, on average.
- For every one point increase in the critics score, the audience score goes up by 0.519 points.
- For every one point increase in the audience score, the critics score goes up by 0.519 points, on average.
:::

:::
::::

::: {.column width="30%"}
{{< include _wooclap-column.qmd >}}
:::
::::::

## Participate 📱💻 {.xsmall}

:::::: columns
:::: {.column width="70%"}

::: wooclap
The intercept of the model for predicting audience score from critics score is 32.3.
Which of the following is the best interpretation of this value?

$$\widehat{\text{audience}} = 32.3 + 0.519 \times \text{critics}$$

::: wooclap-choices
- For movies with a critics score of 0 points, we expect the audience score to be 32.3 points, on average.
- For movies with an audience score of 0 points, we expect the critics score to be 32.3 points, on average.
- For every one point increase in the critics score, the audience score goes up by 32.3 points.
- For movies with an audience score of 0 points, we expect the critics score to be 0.519 points, on average.
:::

:::
::::

::: {.column width="30%"}
{{< include _wooclap-column.qmd >}}
:::
::::::

## Is the intercept meaningful?

✅ The intercept is meaningful in context of the data if

-   the predictor can feasibly take values equal to or near zero or
-   the predictor has values near zero in the observed data

. . .

🛑 Otherwise, it might not be meaningful!

# Application exercise

## `{r} todays_ae` {.smaller}

::: appex
-   Go to your ae project in RStudio.

-   If you haven't yet done so, make sure all of your changes up to this point are committed and pushed, i.e., there's nothing left in your Git pane.

-   If you haven't yet done so, click Pull to get today's application exercise file: *`{r} paste0(todays_ae, ".qmd")`*.

-   Work through the application exercise in class, and render, commit, and push your edits.
:::

## Recap

::: task
Calculate the predicted body weights of penguins on Biscoe, Dream, and Torgersen islands *by hand*.
:::

$$
\widehat{body~mass} = 4716 - 1003 \times islandDream - 1010 \times islandTorgersen
$$

. . .

-   Biscoe: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 0 = 4716$

. . .

-   Dream: $\widehat{body~mass} = 4716 - 1003 \times 1 - 1010 \times 0 = 3713$

. . .

-   Torgersen: $\widehat{body~mass} = 4716 - 1003 \times 0 - 1010 \times 1 = 3706$

## Models with categorical predictors

::: incremental
-   When the categorical predictor has many levels, they're encoded to **dummy variables**.

-   The first level of the categorical variable is the baseline level.
    In a model with one categorical predictor, the intercept is the predicted value of the outcome for the baseline level (x = 0).

-   Each slope coefficient describes the difference between the predicted value of the outcome for that level of the categorical variable compared to the baseline level.
:::
